Everthing you wanted to know about PanDA but were truly truly afraid to ask

### 2016-11-08 bckground.tex
Still issues with capitalizations
Iteration background, preliminary work on workflow

The lack of a consistent and shared definition of ``workflow'' hinders the
understanding of its properties and its relations with related concepts. For
example, we need to clarify the difference among ``workflow'', ``workload'',
``task'', or ``job'' but also between workflow template and instance, or
data-flow and control-flow. This is precondition to specify properly the design
of software systems that support the execution of scientific applications.

@@

Each task may have an arbitrary number of properties like number of cores,
executables, or input/output files. Tasks may have precedence relations
depending on their data dependences or any other type of dependency mandated
by the application algorithms. Tasks with precedence relations have to be
executed serially, otherwise tasks can be executed concurrently. A workload is
defined as a set of tasks that can be executed concurrently. As such, a
workflow can be composed by a set of workloads.

The terms ``task'' and ``job'' are also used inconsistently across communities
that perform scientific computing. In this paper, ``job'' refers to the unit of
work that is submitted to a local resource management system (LRMS), like a
batch system or a scheduler. As such, a task can become a job when is scheduled
on a resource that exposes a LRMS but can also become a virtual machine or a
container when bootstrapped on an infrastructure supporting virtualization.
Tasks can be grouped into jobs, statically or dynamically, depending on the
resource capabilities and the task requirements.

Usually, workflows are represented as graphs in which tasks are vertices and
relations are edges~\cite{}. Often, graphs are supposed to be acyclic but graphs
with cycles have been used to represent workflows of workflows~\cite{}. In this
paper, a workflow template is a type of graph while a workflow instance is a
type of graph with a specific set of vertices and edges. Further, a graph is
``abstract'' when no resource properties are available for all vertices,
``concrete'' otherwise~\cite{}.

The distinction between an abstract and concrete graph highlights an important
element of the execution process of a workflow. Workflows can be specified
without indicating on what resources they should be executed. Accordingly, an
abstract graph can be interpreted as a formal description of the requirements
for the workflow execution. Once these requirements are matched by a set of
resource capabilities, the graph becomes concrete, i.e., a formal description
of the workflow execution process and its execution environment.

>>>>

The lack of a consistent and shared definition of ``workflow'' hinders the
understanding of its properties and its relations with related concepts. For
example, we need to clarify the difference among ``workflow'', ``workload'',
``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
``data-flow'' and ``control-flow''. This is precondition to specify properly the
design of software systems that support the execution of scientific
applications.

@@

Each task may have an arbitrary number of properties like number of cores,
executables, or input/output files. Tasks may have precedence interrelations,
depending on their data dependences or any other type of dependency mandated
by the application algorithms. Tasks with precedence relations have to be
executed serially; otherwise, tasks can be executed concurrently. A workload is
defined as a set of tasks that can be executed concurrently. As such, a
workflow can be composed of a set of workloads.

The terms ``task'' and ``job'' are also used inconsistently across communities
that perform scientific computing. In this paper, ``job'' refers to the unit of
work that is submitted to a local resource management system (LRMS), like the
Slurm or PBS batch system of a cluster. As such, a task can become a job when
is scheduled on a resource that exposes a LRMS but can also become a virtual
machine or a container when bootstrapped on an infrastructure supporting
virtualization. Tasks can be statically or dynamically grouped into jobs,
depending on the resource capabilities and the task requirements.

Usually, workflows are represented as graphs in which tasks are vertices and
relations are edges~\cite{}. Often, graphs are supposed to be acyclic but graphs
with cycles have been used to represent workflows of workflows~\cite{}. In this
paper, a workflow template is a type of graph while a workflow instance is a
workflow template with specific vertices and edges. Further, a workflow instance
is ``abstract'' when no resource properties are available for all vertices,
``concrete'' otherwise~\cite{}. When not qualified, the term ``workflow''
indicates a abstract workflow instance.

The distinction between an abstract and concrete workflow instances highlights
an important element of the execution process of a workflow. Abstract workflows
can be interpreted as a formal description of the requirements for the workflow
execution, while concrete workflows can be interpreted as a formal description of the workflow's execution process. The process of resource selection matches the resource capabilities to the workflow requirements making possible the transition from execution requirements to execution process.

Workload managers are those components of workflow systems or, more in general, of distributed applications that enable resource selection.

\subsection{ATLAS Workflow}

In this paper we focus on the ATLAS Monte Carlo workflow\mtnote{Are there other
workflows for ATLAS?} and its evolving processes of resource selection.
Traditionally, the ATLAS workflow has been based on single-core tasks, executed
on Grid resources to process large amount (2TB) of input data divided into
discrete I/O units called ``events''. This approach has many merits: (i)
simplified process of resource selection under the assumption of homogeneous
resources and execution environments; (ii) opportunistic distribution of compute
tasks across resources based on contingent availability; (iii) relative
robustness due to the ability of re-executing failed tasks; and (iv)
multi-tiered distribution and storage of events' data.

Despite all the advantages,

\begin{figure}
  \includegraphics[width=\columnwidth]{figures/atlas_workflow.pdf}
  \caption{ATLAS Workflow}
\label{fig:atlas_workflow}
\end{figure}
